---
title: "MADONA - TBI outcome prediction"
author: "Florian Pierre MARTIN"
date: "`r Sys.Date()`"
toc: true
number-sections: true
format: 
  html:
    code-fold: true
    df-print: paged
editor: visual
---

```{r}
library(tidymodels)
library(tidyverse)
library(gmodels)
library(naniar)
library(lubridate)
library(tableone)
library(gridExtra)
library(hms)
library(readxl)
library(corrplot)
library(ggplot2)
library(Boruta)
library(mice)
library(missForest)
library(Rcpp)
library(ggpubr)
library(glmnet)
library(ranger)
library(xgboost)
library(keras)
library(discrim)
library(kknn)
library(kernlab)
library(themis)
library(viridis)
library(FactoMineR)
library(factoextra)
library(tableone)
library(FFTrees)
library(ggpubr)
library(probably)
library(betacal)
library(pROC)
library(VIM)
```

# Data preprocessing

## Missing data

### Pattern

```{r}
mice_plot <- aggr(madona_mother %>% dplyr::select(pic_moy, pic_var, evo_scan, evo_scan_bin, starts_with("evl_")), col=c('navyblue','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(madona_mother %>% dplyr::select(pic_moy, pic_var, evo_scan, evo_scan_bin, starts_with("evl_"))), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))

gg_miss_upset(madona_mother)

ggplot(madona_mother,
       aes(x = pic_moy,
           y = evo_scan)) +
 geom_miss_point() + facet_wrap(~evo_scan)

gg_miss_fct(x = madona_mother %>% dplyr::select(pic_moy, pic_var, evo_scan, starts_with("evl_"), glas), fct = glas)
```

### Imputation

#### missForest

```{r}
########### using missForest
set.seed(261)
madona_moth_missfo <- missForest(madona_mother %>% as.data.frame(), variablewise = FALSE, verbose = TRUE, mtry = 20, maxiter = 20, ntree = 200)
madona_moth_missfo$OOBerror
madona_moth_imp <- madona_moth_missfo$ximp
```

#### MICE (pmm)

```{r}
madona_mice <- mice(madona_mother, m = 5, method = if_else(colnames(madona_mother) == "pic_moy" | colnames(madona_mother) == "pic_var", "pmm", if_else(colnames(madona_mother) == "evo_scan_bin" | colnames(madona_mother) == "evl_s" | colnames(madona_mother) == "evl_azi" | colnames(madona_mother) == "evl_ao" | colnames(madona_mother) == "evl_ame" | colnames(madona_mother) == "evl_am1" | colnames(madona_mother) == "evl_am", "logreg", if_else(colnames(madona_mother) == "evo_scan", "polyreg", ""))), seed = 23)

madona_imp_mice <- madona_mice %>%
  complete(action = 1) %>% as_tibble()
```

# Data exploration

## FAMD

```{r}
## show mortality bias
res.famd <- FAMD(madona_moth_imp, graph = FALSE, sup.var = c(38, 39, 51, 55))
fviz_mfa_ind(res.famd, 
             habillage = "deces_m6",
             addEllipses = T, ellipse.type = "euclid",
             geom = "point",
             mean.point = FALSE,
             palette = c("#FAB248", "#91501B"),
             repel = F
) 
ind <- get_famd_ind(res.famd)
ind
coord_famd <- ind$coord %>% cbind(madona_moth_imp$deces_m6) %>% cbind(madona_moth_imp$gose_m6_bin) %>% as.data.frame() %>% dplyr::mutate(V6 = as.factor(V6), V7 = as.factor(V7))
coord_famd %>% ggplot(aes(x = V6, y = Dim.1)) + geom_boxplot(fill = c("#FAB248", "#91501B"), color = "grey") + stat_compare_means() + theme_bw() + coord_flip()
coord_famd %>% ggplot(aes(x = V6, y = Dim.2)) + geom_boxplot(fill = c("#FAB248", "#91501B"), color = "grey") + stat_compare_means() + theme_bw()
# Contribution to the first dimension
fviz_contrib(res.famd, "var", axes = 1, fill = "darkseagreen", color = "white")
```

# Modelling workflow

## Data splitting

```{r}
set.seed(241)
split_a <- initial_split(madona_moth_imp, prop = 0.70, strata = "gose_m6_bin")
training_a <- split_a %>% training()
test_a <- split_a %>% testing()
```

## Resampling - cross-validation

```{r}
set.seed(249)
cv_folds_a <- vfold_cv(training_a, v = 3, repeats = 20, strata = "gose_m6_bin")
```

## Feature selection - Boruta

```{r}
set.seed(244)
boruta_a <- Boruta(x = training_a[, -which(colnames(training_a) == "gose_m6_bin")],
                         y = training_a$gose_m6_bin, doTrace = T, maxRuns = 1000)
getSelectedAttributes(boruta_a, withTentative = F)

plot(boruta_a, xlab = "", xaxt = "n")
lz<-lapply(1:ncol(boruta_a$ImpHistory),function(i)
        boruta_a$ImpHistory[is.finite(boruta_a$ImpHistory[,i]),i])
names(lz) <- colnames(boruta_a$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
     at = 1:ncol(boruta_a$ImpHistory), cex.axis = 0.7)
```

## Recipe

```{r}
recipe_a <- recipe(gose_m6_bin ~ age + evl_ame + glas + craniec_ini + mydriase_bin + evl_azi + evo_scan_bin + pic_var + pic_moy, 
                        data = training_a) %>%
  step_normalize(all_numeric()) %>% 
  step_dummy(all_factor_predictors())
```

## Parsnip

```{r}
logistic_model <- logistic_reg() %>%
        set_mode("classification") %>%
        set_engine("glm")
rf_model <- rand_forest(trees = 1000) %>%
        set_engine("ranger") %>%
        set_mode("classification")
knn_model <- nearest_neighbor(neighbors = 20) %>%
        set_engine("kknn") %>%
        set_mode("classification")
svm_model <- svm_poly() %>%
        set_engine("kernlab") %>%
        set_mode("classification")
```

## Workflow

```{r}
log_wflow_a <- workflow() %>%
        add_model(logistic_model) %>%
        add_recipe(recipe_a)
rf_wflow_a <- workflow() %>%
        add_model(rf_model) %>%
        add_recipe(recipe_a)
knn_wflow_a <- workflow() %>%
        add_model(knn_model) %>%
        add_recipe(recipe_a)
svm_wflow_a <- workflow() %>%
        add_model(svm_model) %>%
        add_recipe(recipe_a)
```

## Hyperparameter tuning

### Tuning models

```{r}
tune_rf_train <-
        rand_forest(
                mtry = tune(),
                trees = 1000,
                min_n = tune(),
        ) %>% 
        set_mode("classification") %>% 
        set_engine("ranger")

tune_knn_train <-
        nearest_neighbor(
                neighbors = tune()) %>% 
        set_mode("classification") %>% 
        set_engine("kknn")

tune_svm_train <-
        svm_poly(cost = tune()) %>% 
        set_mode("classification") %>% 
        set_engine("kernlab")
```

### Random Forest

```{r}
tune_rf_wf_a <-
        workflow() %>% 
        add_recipe(recipe_a) %>% 
        add_model(tune_rf_train)

set.seed(529)
tune_rf_res_a <-
        tune_grid(
                tune_rf_wf_a,
                resamples = cv_folds_a,
                grid = 15,
                metrics = metric_set(roc_auc)
        )
tune_rf_res_a %>%
        collect_metrics() %>%
        filter(.metric == "roc_auc") %>%
        select(mean, min_n, mtry) %>%
        pivot_longer(min_n:mtry,
                     values_to = "value",
                     names_to = "parameter"
        ) %>%
        ggplot(aes(value, mean, color = parameter)) +
        geom_point(show.legend = FALSE) +
        facet_wrap(~parameter, scales = "free_x") +
        labs(x = NULL, y = "ROC AUC")
rf_grid_a <- grid_regular(
        mtry(range = c(1, 3)), 
        min_n(range = c(15, 25)),
        levels = 10
)
set.seed(529)
tuned_train_rf_res_a <- tune_grid(
        tune_rf_wf_a,
        resamples = cv_folds_a,
        grid = rf_grid_a, 
        metrics = metric_set(roc_auc, kap, brier_class), control = control_resamples(save_pred = TRUE)
)
tuned_train_rf_res_a %>%
        collect_metrics() %>%
        filter(.metric == "roc_auc") %>%
        mutate(min_n = factor(min_n)) %>%
        ggplot(aes(mtry, mean, color = min_n)) +
        geom_line(alpha = 0.5, linewidth = 1.5) +
        geom_point() +
        labs(y = "ROC AUC")
```

### K nearest neighbours

### Support vector machine

### Best algorithm (for each model)

```{r}
############## RF
best_rf_roc_a <- select_best(tuned_train_rf_res_a, "roc_auc")

final_rf_a <-
        finalize_model(
                tune_rf_train,
                best_rf_roc_a
        )
final_rf_wf_a <- workflow() %>%
        add_recipe(recipe_a) %>%
        add_model(final_rf_a)
```

### Best model selection

```{r}
auc_rf_train_a <- 
        collect_metrics(tuned_train_rf_res_a, summarize = F) %>%
        filter(.metric == "roc_auc") %>%
        select(.estimate) %>%
        pull() %>%
        sort()

auc_df_rf_train_a <- data.frame(auc_rf_train_a) %>%
        mutate(models = "RF") %>%
        mutate(auc_res = auc_rf_train_a)
```

### Bootstrap

```{r}
set.seed(255)
testset_boot_a <- bootstraps(test_a, times = 1000, strata = "gose_m6_bin") 

class_metric <- metric_set(roc_auc, brier_class, kap, f_meas, sensitivity, specificity, accuracy)

############### BOOTSTRAP SUPPORT VECTOR MACHINE
svm_fit_boot_a <-
        final_svm_wf_a %>% 
        fit_resamples(testset_boot_a, metrics = class_metric, control = control_grid(save_pred = TRUE))

auc_test_svm_a <- 
        collect_metrics(svm_fit_boot_a, summarize = F) %>%
        filter(.metric == class_metric) %>%
        select(.estimate) %>%
        pull() %>%
        sort()
mean(auc_test_svm_a)
quantile(auc_test_svm_a, probs = c(0.25, 0.75))
```

# Model performances

```{r}
##################### MODEL A
# ROC curve
svm_fit_boot_a %>% collect_predictions() %>% roc_curve(gose_m6_bin, .pred_0) %>% ggplot() + geom_line(aes(x = 1-specificity, y = sensitivity), color = "#FC8D62", linewidth = 1.5) + geom_abline(intercept = 0, slope = 1, linetype = "dotted") + theme_bw()
# Se for Sp 95
svm_fit_boot_a %>% collect_predictions() %>% roc_curve(gose_m6_bin, .pred_0) %>% dplyr::filter(specificity > 0.949 & specificity < 0.951) %>% dplyr::select(sensitivity)
# performance metrics
auc_test_svm_a_df <- data.frame(auc_test_svm_a) %>%
        mutate(metric_value = "AUROC") %>%
        mutate(perf_res = auc_test_svm_a)

###### PPV & NPV
# prevalence of the positive event --> we fixed the prevalence at 50%
madona_moth_imp %>% dplyr::select(gose_m6_bin) %>% base::table() %>% prop.table()
# PPV
svm_fit_boot_a %>% collect_predictions(summarize = F) %>% ppv(truth = gose_m6_bin, estimate = .pred_class, prevalence = 0.45)

dat_ppv_a <- data.frame(
  category = c("PPV", ""), 
  ppv_val = c(svm_fit_boot_a %>% collect_predictions() %>% ppv(truth = gose_m6_bin, estimate = .pred_class, prevalence = 0.45) %>% dplyr::select(.estimate) %>% pull(), 1-svm_fit_boot_a %>% collect_predictions() %>% ppv(truth = gose_m6_bin, estimate = .pred_class, prevalence = 0.45) %>% dplyr::select(.estimate) %>% pull())
)
dat_ppv_a$ymax <- cumsum(dat_ppv_a$ppv_val)
dat_ppv_a$ymin <- c(0, head(dat_ppv_a$ymax, n = -1))
dat_ppv_a %>% ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = category)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  scale_fill_manual(values = c("lightgrey", "#FC8D62")) +
  theme_void()

# NPV
svm_fit_boot_a %>% collect_predictions(summarize = F) %>% npv(truth = gose_m6_bin, estimate = .pred_class, prevalence = 0.45)

dat_npv_a <- data.frame(
  category = c("NPV", ""), 
  npv_val = c(svm_fit_boot_a %>% collect_predictions() %>% npv(truth = gose_m6_bin, estimate = .pred_class, prevalence = 0.45) %>% dplyr::select(.estimate) %>% pull(), 1-svm_fit_boot_a %>% collect_predictions() %>% npv(truth = gose_m6_bin, estimate = .pred_class, prevalence = 0.45) %>% dplyr::select(.estimate) %>% pull())
)
dat_npv_a$ymax <- cumsum(dat_npv_a$npv_val)
dat_npv_a$ymin <- c(0, head(dat_npv_a$ymax, n = -1))
dat_npv_a %>% ggplot(aes(ymax = ymax, ymin = ymin, xmax = 4, xmin = 3, fill = category)) +
  geom_rect() +
  coord_polar(theta = "y") +
  xlim(c(2, 4)) +
  scale_fill_manual(values = c("lightgrey", "#FC8D62")) +
  theme_void()
```
